---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Introduction to centraldogma

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experinental)
<!-- badges: end -->

```{r setup, include = FALSE}
library(centraldogma)
```

The [centraldogma](https://github.com/rforbiodatascience22/group_10_package)
package replicates the central dogma of molecular biology. More specifically it
replicates the flow of genetic information from DNA to RNA to protein.

The package consists of the following five functions, which are demonstrated
below: `generate_dna`, `transcribe`, `split_codons`, `translate` and
`plot_aa_occurrence`.

### Generate DNA

The `generate_dna` is used to generate a random DNA sequence. The `length`
argument controls the length of the generated sequence:

```{r}
set.seed(6942069)

generate_dna(length = 20)
```


### Transcribe DNA to RNA

The `transcribe()` function transcribes a DNA sequence into a RNA sequence by 
replacing Thymine (T) with Uracil (U):

```{r}
dna_fragment <- "CCATGTTATG"

transcribe(dna_fragment)
```


### Split RNA into Codons

The `split_codons` function splits a RNA sequence into codons. The `start`
argument specifies the start position of the reading frame. The position can be
any position in the RNA sequence with a default value of one:

```{r}
rna_fragment <- "AUCGUACGAUAUGAUACAGAGAUAGACAUAUUUAACGG"

split_codons(rna_fragment, start = 5)
```


### Translate RNA Codons into Protein

The `translate` function translates a sequence of RNA codons into a protein
sequence by replacing codons into the amino acids they code for using a standard
codon table. Stop codons are translated to "_". The codon table can be accessed
with the function `codon_table`.

Here is an example of the use of the `translate` function:

```{r}
rna_codon_sequence <- c("GGG", "GGC", "GCG", "UCC", "GUC")

translate(rna_codon_sequence)
```


### Plot Amino Acid Occurrences in a Protein 

The `plot_aa_occurrence` function makes a plot of the occurrence of each amino
acid in a protein sequence:

```{r, fig.width=6, fig.height=2}
protein_sequence <- "GGASVTVSRFW_PSQSKQRHRVEPVS_IQSYLP"

plot_aa_occurrence(protein_sequence)
```


### Example Workflow

The five functions included in `centraldogma` can be used in a pipeline as
follows:

```{r, eval = FALSE}
dna_sequence <- generate_dna(length = 250)
rna_sequence <- transcribe(dna_sequence)
rna_codon_sequence <- split_codons(rna_sequence)
protein_sequence <- translate(rna_codon_sequence)
plot_aa_occurrence(protein_sequence)
```


### Use cases for package

`centraldogma` can e.g. be used to translate DNA sequences into protein
sequences by running the full pipeline of functions as shown under example
workflow. Alternatively one could start with RNA sequences instead and do the
translation or start with a protein sequence and make an occurrence plot.

### Ideas for more functions for the package

Some other functions that could be implemented in the package could be:

* A function to read FASTA-files
* A function to write FASTA-files
* A function for reverse transcription

Also it would be nice if the `translate` function could use other codon tables
than the standard one.

### Main points from discussion in Task 4

The number of dependencies should be limited such that the user does not have to
install a bunch of packages. It cannot be avoided if the R base package does not
include the functionality that you want and you do not wish to implement it
yourself.

Adding an `@importFrom package function` tag to a function description
has a small performance benefit compared to `package::function()` due to `::`
adding approximately 5 µs to function evaluation time. However, it is more clear
what package a function belongs to with `::`. Also, `@importFrom` might cause
name conflicts if another function with the same name already exists in the
namespace.
